---
output:
  github_document:
    html_preview: true
---

# Notes for build
  * use precompiled .Rmd for vignettes ([a la](https://ropensci.org/blog/2019/12/08/precompute-vignettes/)) because of build fails on win-builder.
    - knit tutorial with `knitr::knit("vignettes/tutorial.Rmd.orig", output = "vignettes/tutorial.Rmd")`
    - add *.Rmd.orig to Rbuildignore
    - update and reknit on new submission/commit/release
    -set knit to `fig.path=""`
    - manually move figures from trackter root to vignette with code below
    - add r code for users with `knitr::purl("vignettes/tutorial.Rmd.orig", output = "vignettes/tutorial.R")`

```{r, compile vignette, eval=FALSE}

knitr::knit("vignettes/tutorial.Rmd.orig", output = "vignettes/tutorial.Rmd")

knitr::purl("vignettes/tutorial.Rmd.orig", output = "vignettes/tutorial.R")
```

```{r, eval=FALSE}
f = list.files(full.names = TRUE,pattern = ".png")
f2 <- gsub("./","./vignettes/",f)
file.copy(f,f2)
sapply(f,unlink)
``` 
# Notes for testing/CI

    * follow these pointers for `pkgdown` site build with CI on Travis
    * check travis.yaml with https://config.travis-ci.com/explore
    * imagemagick requirements related to ubuntu/linux https://srvanderplas.netlify.app/post/2019-05-07-travis-tricks/
    
# Notes for gitting/site

  * `usethis::use_pkgdown()` for `pkgdown`
  * `pkgdown::build_site()` to build site
  * see: https://cran.r-project.org/web/packages/pkgdown/vignettes/pkgdown.html
  



 