---
output:
  github_document:
    html_preview: true
---
[![Build Status](https://travis-ci.com/ckenaley/trackter.svg?branch=master)](https://travis-ci.com/ckenaley/trackter) [![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ckenaley/trackter?branch=master&svg=true)](https://ci.appveyor.com/project/ckenaley/trackter) [![](https://www.r-pkg.org/badges/version/trackter?color=orange)](https://cran.r-project.org/package=trackter)[![Codecov test coverage](https://codecov.io/gh/ckenaley/trackter/branch/master/graph/badge.svg)](https://codecov.io/gh/ckenaley/trackter?branch=master)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
library(trackter)
library(ggplot2)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figure/",
  fig.height = 1
  
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# *trackter*

## Description

*trackter* is an R package for semiautomated tracking and analysis of 2D kinematics from video and image data. The core functions of *trackter* automatically detect a region of interest (ROI) and compute important kinematic and shape parameters based on the ROI's contour. These functions use thresholding and segmentation to identify ROIs and, thus, moderately contrasted images are required. Presented below is some of *trackter*'s functionality in the context of fish locomotion.


Please report any bugs or performance issues.

## Getting started

Please check out *trackter*'s [github page](https://ckenaley.github.io/trackter/)


## Installation

The release version of *trackter* can be installed with:

```{r eval=FALSE}
install.packages("trackter")
```

The current development version can be installed with:

```{r eval=FALSE}
    require(devtools)
    install_github("ckenaley/trackter")
    require(trackter)
```
    


## External dependencies

The core functions of *trackter* that extract shape and contour data from images ( `kin.simple` and `kin.search`) depend upon *EBImage*, available on the Bioconductor repository. The current build and development versions of *trackter* install this dependency. If it does not install, it can be done so easily with just a few lines of code: 

```{r eval=FALSE}
  if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
   BiocManager::install("EBImage")
```

*trackter* also contains several functions for image and video processing. These functions depend on the popular `FFmpeg` package and it must be installed if the user intends to use them. Installation is platform-dependent. I found the [`FFmpeg` wiki installation and compilation guide to be quite useful](https://trac.ffmpeg.org/wiki/CompilationGuide).

## Features

__Automated kinematic analysis__

* Fast and accurate contour and shape analysis of ROIs.
* ROI detection with search parameters including position and size.
* Relevant functions: `kin.search` and `kin.simple`.

__Tools for kinematic analysis of swimming animals__

* Calculate midline (propulsive) wavelength, trailing-edge frequency, paired-fin position.
* Relevant functions: `amp.freq`, `halfwave`, `wave`, and `fin.kin`

__Tools for image and video processing using FFmpeg__

* Access `FFmpeg` functionality, including filters and codecs, to extract frames, stitch videos, and edit images and videos.

* Relevant functions: `images.to.videos`, `images.to.videos2`, `vid.to.images`, and `vid.to.images2`

__Other miscellaneous, low-level tools for kinematic analysis__

* Compute distances in 2d space, angles, heading/bearing, convert radians to degrees and vice versa.
* Relevant functions: `dist.2d`, `cosine.ang`, `bearing.xy`, `deg`, `rad`
